---
title: "<span style='color:darkblue;'>A Global Perspective:</span>"
subtitle: "<span style='color:darkblue;'>Navigating the Weight of Tommorow, Today</span>" 
author: Aoife Kelly
date: "2024-04-01"
format: 
  html:
    theme: sandstone
    backgroundcolor: lightblue
execute: 
  echo: false
  warning: false
  message: false
---

## Introduction

### Critical Global Health Issue regarding Overweight Children

The amount of children under the age of 5 who are statistically <span style="color:red;">overweight</span> is becoming a critical global health issue impacting the well-being of children worldwide.The prevalence of <span style="color:blue;">overweight</span> children in this age group has been steadily increasing, posing significant challenges to their physical health and future development. As we embark on this journey of <span style="color:blue;">**'Navigating the Weight of Tommorow, Today'**</span>, we must delve into the depths of <span style="color:purple;">***UNICEF***</span> data to unravel the story of our youngest generation grappling with the weight of today's lifestyle choices.

This interactive dashboard offers a comprehensive examination of childhood <span style="color:red;">overweight</span> rates across different regions of the world, drawing on data from <span style="color:purple;">***UNICEF***</span>. Through a series of visualizations and insights, we aim to:

1. **Illuminate the prevalence of overweight children**
2. **Identify geographical variances**
3. **Underscore the urgency of addressing this growing concern**

It is important to engage with the data to gain deeper insights into the factors contributing to childhood overweight in order *to foster healthier environments and lifestyles for our youngest population.*


```{r}
#| label: setup
#| echo: false

#libraries to load
library(tidyverse) 
library(plotly)

# data
unicef_metadata <- read_csv("unicef_metadata.csv", 
                            col_types = cols(alpha_2_code = col_skip(), 
                                             alpha_3_code = col_skip(), numeric_code = col_skip(), 
                                             iso3c = col_skip()))
unicef_indicator_1 <- read_csv("unicef_indicator_1.csv", 
                               col_types = cols(alpha_2_code = col_skip(), 
                                                alpha_3_code = col_skip(), numeric_code = col_skip(), 
                                                indicator = col_skip(), sex = col_skip(), 
                                                unit_multiplier = col_skip(), unit_of_measure = col_skip(), 
                                                observation_status = col_skip(), 
                                                observation_confidentaility = col_skip(), 
                                                time_period_activity_related_to_when_the_data_are_collected = col_skip(), 
                                                current_age = col_skip()))
data_right_3 <- read_csv("data_right_3.csv")


# transformations
data_join <- full_join(unicef_metadata, unicef_indicator_1)
data_join <- full_join(unicef_metadata,unicef_indicator_1, by = c("country", "year" = "time_period"))
data_join <- full_join(unicef_indicator_1, data_right_3)
data_join <- full_join(unicef_indicator_1, data_right_3, by = c("country"))
data_join <- full_join(unicef_metadata, data_right_3)
data_join <- full_join(unicef_metadata, data_right_3, by = c("country"))

# Final Data Object
data_join <- unicef_metadata %>%
  full_join(unicef_indicator_1, by = c("country", "year" = "time_period")) %>%
  full_join(data_right_3, by = c("country"))

```

```{r}
#| fig-cap: "The two world maps help depict the number of overweight children (under 5) per country in 2000 and 2022 by providing a visual representation of the changes in the childhood overweight category over two decades. In order to show comparison overtime the data in put in map format for both 2000 and 2022, viewers can quickly compare and contrast the prevalence of childhood obesity in different countries across the two time periods. This visual comparison enables easy identification of trends and changes over time. With the help of colour-coding to represent data, where darker shades indicate higher numbers or percentages. In this case, viewers can easily see which countries had higher levels of childhood overweightness in each year and how those levels have changed over time. For instance, if a country's shade becomes lighter from 2000 to 2022, it suggests a decrease in childhood overweightness, whereas a darker shade indicates an increase.Countries flagged by the visualisation :From 2000 to 2022, Russia's concerted efforts saw a remarkable decline in childhood obesity, with numbers halved from 1,051,300 to 559,500 children overweight (under 5). Contrastingly, Indonesia's struggle against childhood obesity intensified, witnessing a stark increase from 694,000 to 2,357,600 overweight children (under 5), highlighting pressing health challenges."

# Map 1 
map_world <- map_data("world")

data_join %>%
  filter(year == 2022) %>%
  full_join(map_world, by = c("country" = "region")) %>%
  ggplot(aes(x = long, y = lat, group = group, fill = obs_value)) +
  geom_polygon() +
  scale_fill_gradient(low = "red", high = "purple", na.value = "lightgrey") +
  labs(
    title = "Differences between countries regarding the amount of Overweight Children in 2000",
    subtitle = "Countries in grey have no data due to a mismatch with their names",
    caption = "Source : R package {gapminder}",
    x = "Longitude",
    y = "Latitude",
    fill = "obs_value"
  ) +
  theme_bw()

```

```{r}
# Map 2 
map_world <- map_data("world")

data_join %>%
  filter(year == 2022) %>%
  full_join(map_world, by = c("country" = "region")) %>%
  ggplot(aes(x = long, y = lat, group = group, fill = obs_value)) +
  geom_polygon()

```


